---
title: "SIR metapop"
output: pdf_document
date: "2024-12-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      fig.align = "center",
                      message = FALSE,
                      warning = FALSE,
                      results = 'show')

knitr::opts_knit$set(root.dir='C:/Users/wbenhamou/Desktop/OMEGA/Data_and_codes')
```
\newpage

# Initialization

```{r, results='hide'}
# Cleaning objects from the workplace
rm(list=ls())

# Packages (may first require installations: install.packages())
library(tidyverse) # for data manipulation
library(tidyr) # for data manipulation
library(scales) # to manipulate the internal scaling infrastructure used by ggplot2
library(RColorBrewer) # for colors and palettes
library(viridisLite)
library(gridExtra) # to show multiple plots
library(cowplot) # to show multiple plots
library(ggpubr) # to show multiple plots
library(deSolve) # to solve a system of ordinary differential equations
library(boot) # for the inverse logit function 'inv.logit'
library(rootSolve)
library(ggpattern)
library(ggnewscale)

# Execute functions from external R script
source("Functions.R")

Sys.setlocale("LC_TIME", "English")
```

# Evolutionary dynamics of an emerging variant in a two-patch metapopulation.

```{r}
nA <- 100
nB <- 100

q0 <- 1/100

I0w <- 1E-7
I0m <- q0/(1-q0)*I0w

init <- c("SA"=nA-I0w-I0m, "IAw"=I0w, "IAm"=I0m, "RA"=0,
          "SB"=nB-I0w, "IBw"=I0w, "IBm"=0, "RB"=0)

tf <- 255
tM <- 80
times <- seq(0,tf,1)

x_breaks <- seq(-50,tf-tM,50)

bAw <- bBw <- 0.15
bAm <- bBm <- bBw*1.3 # Higher transmission rate

gAw <- gBw <- 0.1
gAm <- gBm <- 0.1

mA <- 0
mB <- 0.1

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
           "mA"=mA, "mB"=mB)

simul <- simulate2(init, times=times, parms=parms, tM=tM, qB0=q0)
simul$time <- simul$time-tM
```

```{r, fig.width=8, fig.height=4}
mA_vec <- ifelse(simul$time<0, yes=0, no=mA)
mB_vec <- ifelse(simul$time<0, yes=0, no=mB)

simul$sA <- (bAm-bAw)*simul$SA/nA-(gAm-gAw)+
  (mA_vec*((bBm-bBw)*(simul$IBm/nB)/(simul$IAm/nA)-(bAm-bAw))+
     nB/nA*mB_vec*(bAm-bAw)*((simul$IBm/nB)/(simul$IAm/nA)-1))*simul$SA/nA

simul$sB <- (bBm-bBw)*simul$SB/nB-(gBm-gBw)+
  (mB_vec*((bAm-bAw)*(simul$IAm/nA)/(simul$IBm/nB)-(bBm-bBw))+
     nA/nB*mA_vec*(bBm-bBw)*((simul$IAm/nA)/(simul$IBm/nB)-1))*simul$SB/nB

simul$hA <- -(simul$qA-simul$qB)*(mA_vec*bBw+nB/nA*mB_vec*bAw)*(simul$IBw+simul$IBm)/(simul$IAw+simul$IAm)*
  simul$SA/nB/(simul$qA*(1-simul$qA))
simul$hB <- -(simul$qB-simul$qA)*(mB_vec*bAw+nA/nB*mA_vec*bBw)*(simul$IAw+simul$IAm)/(simul$IBw+simul$IBm)*
  simul$SB/nA/(simul$qB*(1-simul$qB))

simul$slopeA <- simul$sA+simul$hA
simul$slopeB <- simul$sB+simul$hB

tab <- simul %>% subset(Approx_WM==FALSE) %>% select(time, sB, hB)
tab2 <- data.frame()

for(i in 1:nrow(tab)){
  tab2 <- tab2 %>%
    rbind(data.frame("time"=tab$time[i], "lower"=c(0, tab$hB[i]),
                     "upper"=c(tab$hB[i], tab$sB[i]+tab$hB[i]), "id"=c("hB", "sB")))
}
tab2$id <- factor(tab2$id, levels=c("sB", "hB"))

colors <- c('#da6967', '#4c97cc')

plot_grid(
  
  simul[, colnames(simul) %in% c("time", "logit_qA", "logit_qB", "Approx_WM")] %>%
    gather(variable, value, -time, -Approx_WM) %>%
    subset(Approx_WM==FALSE) %>%
    ggplot(aes(x=time, y=value, col=variable, lty=Approx_WM)) +
      geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
      geom_line(lwd=0.75) +
      labs(x='\nTime', y='Logit-frequency of the variant\n', col="Population") +
      scale_color_manual(values=colors, labels=c("A","B")) +
      scale_x_continuous(breaks=x_breaks, expand=c(0,0)) +
      scale_y_continuous(expand=c(0,0), limits=c(logit(q0),7), breaks=seq(-5,5,2.5)) +
      scale_linetype_manual(values=c(1,3)) +
      annotate(geom='text', label=paste0("Variant: +", (bAm/bAw-1)*100, "% transmission rate"),
               x=-70, y=6.2, hjust=0, size=3.5, fontface='italic') +
      theme_classic() +
      theme(axis.title.x=element_text(size=12), axis.text.x=element_text(size=10),
            axis.title.y=element_text(size=12), axis.text.y=element_text(size=11),
            legend.position=c(0.16,0.71), legend.background=element_rect(colour='black',linewidth=0.2)) +
      guides(lty='none') +
    
    annotation_custom(ggplotGrob(
      simul[, colnames(simul) %in% c("time", "slopeA", "slopeB", "Approx_WM")] %>%
        gather(variable, value, -time, -Approx_WM) %>%
        subset(Approx_WM==FALSE) %>%
        ggplot(aes(x=time, y=value, col=variable, lty=Approx_WM)) +
        geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
        geom_line(lwd=0.4) +labs(x='Time', y='Slope') +
        scale_color_manual(values=c('#da6967', '#4c97cc'), labels=c("A","B")) +
        scale_x_continuous(breaks=seq(-50,150,50), expand=c(0,0)) +
        scale_y_continuous(expand=c(0,0), limits=c(-0.15,0.75), breaks=seq(0,0.8,0.2)) +
        scale_linetype_manual(values=c(1,3)) +
        theme_classic() +
        theme(axis.title.x=element_text(size=8), axis.text.x=element_text(size=7),
              axis.title.y=element_text(size=10), axis.text.y=element_text(size=7),
              legend.position='none')),
    xmin=50, xmax=175, ymin=-4.4, ymax=0.25), # 6 x 4
  
  plot_grid(
    ggplot(subset(simul, Approx_WM==FALSE), aes(x=time)) +
      geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
      geom_line(aes(y=qA-qB), color="black") +
      geom_line(aes(y=(qB*(1-qB))), color=colors[2]) +
      scale_y_continuous(name=expression(q^A-q^B), breaks=seq(0,0.3,0.1), limits=c(0,0.3), expand=c(0,0),
                         sec.axis=sec_axis(~.*1,name=bquote("Genetic variance (pop. B), " ~ q^B * (1 - q^B)))) +
      scale_x_continuous(breaks=x_breaks, expand=c(0,0)) +
      labs(x="Time") +
      theme_classic() +
      theme(axis.title.x=element_blank(),
            axis.text.x=element_text(size=9),
            axis.title.y.left=element_text(size=11),
            axis.text.y.left=element_text(size=8),
            axis.title.y.right=element_text(size=7.5, color=colors[2]),
            axis.text.y.right=element_text(size=8, color=colors[2]),
            axis.ticks.y.right=element_line(color=colors[2]),
            axis.line.y.right=element_line(color=colors[2])),
    
    tab2 %>%
      ggplot(aes(x=time)) +
      geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
      geom_ribbon(aes(ymin=lower, ymax=upper, fill=id)) +
      labs(x="Time", y="Slope (pop. B)") +
      scale_x_continuous(breaks=x_breaks, expand=c(0,0)) +
      scale_y_continuous(expand=c(0,0)) +
      scale_fill_manual(values=c(colors[2],'#99c2e6'), labels=c("S^B", "H^B/(q^B(1-q^B))")) +
      theme_classic() +
      theme(axis.title.x=element_text(size=10), axis.text.x=element_text(size=9),
            axis.title.y=element_text(size=9), axis.text.y=element_text(size=8),
            legend.title=element_blank(), legend.text=element_text(size=7),
            legend.position=c(0.7,0.7)),
    ncol=1, labels = c('B', 'C'), align = 'v'),
  ncol=2, rel_widths = c(0.6,0.4), labels=c('A')) # 8 x 4

# ggsave("Dynamics_logit_freq_AB_transmission_V3.png", width=8, height=4, dpi=1000)
```

```{r, fig.width=4.4, fig.height=5.6}
tf2 <- 160

plot_grid(
  
  subset(simul[seq(1,nrow(simul),4),], Approx_WM==FALSE & time<=tf2) %>%
    ggplot() +
    geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
    geom_abline(aes(slope=slopeA, intercept=logit_qA-slopeA*time, col=time), lwd=0.4) +
    geom_point(aes(x=time, y=logit_qA), cex=0.7) +
    labs(title="Population A", x='\nTime', y='Logit-frequency of the variant\n') +
    scale_color_gradientn(colors=c("#f9d3d2", "#da6967", "#7a1c1b")) +
    scale_x_continuous(limits=c(-tM,tf2), breaks=x_breaks, expand=c(0,0)) +
    scale_y_continuous(expand=c(0,0), limits=c(logit(q0),7), breaks=seq(-5,5,2.5)) +
    theme_classic() +
    theme(axis.title.x=element_blank(), axis.text.x=element_text(size=10),
          axis.title.y=element_text(size=12), axis.text.y=element_text(size=11),
          legend.position='none'),

  subset(simul[seq(1,nrow(simul),4),], Approx_WM==FALSE & time>=0 & time<=tf2) %>%
    ggplot() +
    geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
    geom_abline(aes(slope=slopeB, intercept=logit_qB-slopeB*time, col=time), lwd=0.4) +
    geom_point(aes(x=time, y=logit_qB), cex=0.7) +
    labs(title="Population B", x='\nTime', y='Logit-frequency of the variant\n') +
    scale_color_gradientn(colors = c("#cce5f5", "#4c97cc", "#003f5c")) +
    scale_x_continuous(limits=c(-tM,tf2), breaks=x_breaks, expand=c(0,0)) +
    scale_y_continuous(expand=c(0,0), limits=c(logit(q0),7), breaks=seq(-5,5,2.5)) +
    theme_classic() +
    theme(axis.title.x=element_text(size=12), axis.text.x=element_text(size=10),
          axis.title.y=element_text(size=12), axis.text.y=element_text(size=11),
          legend.position='none'),
  
  ncol=1, rel_heights=c(0.47,0.53))
```

# Evolutionary dynamics of an emerging variant in a host metapopulation with more than two pop.

```{r}
n <- 9

N <- 100

qi0 <- 1/100

I0w <- 1E-7
I0m <- qi0/(1-qi0)*I0w

bAw <- bBw <- 0.11
bAm <- bBm <- bBw*1.15

gAw <- gBw <- 0.1
gAm <- gBm <- 0.1

init <- c("S"=c(N-I0w-I0m, rep(N-I0w,n-1)),
          "Iw"=rep(I0w,n),
          "Im"=c(I0m, rep(0,n-1)),
          "R"=rep(0,n))

tf <- 500
tM <- c(seq(40,40*(n-1),40),tf)

m <- 1E-2
M <- matrix(0, ncol=n, nrow=n)
diag(M) <- 1

parms <- list("bw"=bAw, "bm"=bAm, "gw"=gAw, "gm"=gAm, "M"=M)

simul <- simulate.n(init=init, times=seq(0,tM[1],0.1), parms=parms)

for(i in 2:length(tM)){
  
  parms$M[1:i,1:i] <- c(rep(c(1-m*(i-1), rep(m,i)),i-1), 1-m*(i-1))
  
  init <- unlist(simul[nrow(simul),2:(4*n+1)])
  init[[paste0("Im",i)]] <- qi0/(1-qi0)*init[[paste0("Iw",i)]]
  
  simul <- simul[-nrow(simul),] %>% rbind(simulate.n(init=init, times=seq(tM[i-1],tM[i],0.1), parms=parms))
}

simul[mapply(is.infinite, simul)] <- NA

colors <- c("#da6967", "#e7c459","#5fbb55","#4c97cc",
            "#bc67da","#daa167","#da67ab","#67c3da","#8a96ac")

simul[,c(1, grep("logit_q", colnames(simul)))] %>%
  gather(id, value, -time) %>%
  ggplot(aes(x=time, y=value, col=id)) +
  geom_vline(xintercept=c(0, tM[-length(tM)]), col=colors, lty='dashed', cex=0.3) +
  geom_line(cex=0.7) +
  labs(x="\nTime", y="Logit-frequency of the variant\n", col="Pop.") +
  scale_x_continuous(expand=c(0,0), limits=c(-1,tf)) +
  scale_y_continuous(limits=c(logit(qi0),2.5), breaks=-4:2, expand=c(0,0)) +
  scale_color_manual(values=colors, labels=LETTERS[c(1:n-1,n+1)]) +
  theme_classic() +
  
  annotation_custom(ggplotGrob(
    data.frame("x"=1:n, "id"=rep(1:n,each=n),
               "y"=c(c(NA, rep(1, 8)), 2:9 %>% sapply(function(x){return(c(rep(NA,x-1), rep(x,n-x+1)))})) %>% c) %>%
      ggplot(aes(x=x, y=y, col=as.factor(id))) +
      geom_point(cex=2, alpha=0.8) +
      labs(x="\nConnected populations") +
      scale_color_manual(values=colors) +
      scale_x_continuous(limits=c(2,n)) +
      theme_void() + theme(legend.position='none', axis.title.x=element_text(size=7))),
    xmin=40, xmax=350, ymin=0, ymax=2.5)
  
# 5.5 x 4
```

# The effect of population sizes on the dynamics of the variant frequency in the focal pop.

```{r}
mA <- mB <- 1E-2

prev0 <- 1E-6
qB0 <- 1/100

nA <- 100
nratio_vec <- c(0.02,0.1,1,10,50)

greens <- rev(colorRampPalette(brewer.pal(9,"Greens"))(length(nratio_vec)+2)[-(1:2)])

times <- seq(0,170,0.5)
```

# Higher transmission rate

```{r, fig.width=5, fig.height=4}
bAw <- bBw <- 0.15
bAm <- bBm <- bAw*1.3
gAw <- gBw <- gAm <- gBm <- 0.1

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
           "mA"=mA, "mB"=mB)

simul <- nratio_vec %>% lapply(function(nratio){
  nB <- nratio*nA
  init <- c("SA"=nA*(1-prev0), "IAw"=nA*prev0, "IAm"=0, "RA"=0,
            "SB"=nB*(1-prev0), "IBw"=0, "IBm"=nB*prev0, "RB"=0)
  return(simulate(init, times=times, parms=parms) %>% cbind('n_ratio'=nratio))
}) %>% bind_rows

simul$n_ratio <- factor(simul$n_ratio, levels=sort(nratio_vec, decreasing=TRUE))

Fig2A <- ggplot(simul, aes(x=time, y=logit_qA, col=n_ratio, lty=Approx_WM)) +
  geom_hline(yintercept=0, cex=1, col='grey80') +
  geom_line(lwd=0.7) +
  labs(title="Higher transmission rate\n",
       x='Time', y='Logit-frequency of the variant (pop. A)\n', col=expression(N^B/N^A)) +
  scale_x_continuous(expand=c(0,0), breaks=seq(0,200,50)) +
  scale_y_continuous(expand=c(0,0), limits=c(-7,11), breaks=seq(-5,10,2.5)) +
  scale_color_manual(values=greens) +
  scale_linetype_manual(values=c(1,3)) +
  theme_classic() +
  theme(axis.title.x=element_text(size=10), axis.text.x=element_text(size=10),
        axis.title.y=element_text(size=11), axis.text.y=element_text(size=10),
        legend.title=element_text(size=10), legend.text=element_text(size=10),
        plot.title=element_text(size=11, face='bold.italic', hjust=0.5)) +
  guides(lty='none')
```

# Neutral case (no selection)

```{r, fig.width=5, fig.height=4}
bAw <- bBw <- bAm <- bBm <- 0.15
gAw <- gBw <- gAm <- gBm <- 0.1

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
           "mA"=mA, "mB"=mB)

simul <- nratio_vec %>% lapply(function(nratio){
  nB <- nratio*nA
  init <- c("SA"=nA*(1-prev0), "IAw"=nA*prev0, "IAm"=0, "RA"=0,
            "SB"=nB*(1-prev0), "IBw"=0, "IBm"=nB*prev0, "RB"=0)
  return(simulate(init, times=times, parms=parms) %>% cbind('n_ratio'=nratio))
}) %>% bind_rows

simul$n_ratio <- factor(simul$n_ratio, levels=sort(nratio_vec, decreasing=TRUE))

Fig2B <- ggplot(simul, aes(x=time, y=logit_qA, col=n_ratio, lty=Approx_WM)) +
  geom_hline(yintercept=0, cex=1, col='grey80') +
  geom_line(lwd=0.7) +
  labs(title="No selection\n", x='Time', y='Logit-frequency of the variant (pop. A)\n',
       col=expression(N^B/N^A)) +
  scale_x_continuous(expand=c(0,0), breaks=seq(0,200,50)) +
  scale_y_continuous(expand=c(0,0), limits=c(-7,11), breaks=seq(-5,10,2.5)) +
  scale_color_manual(values=greens) +
  scale_linetype_manual(values=c(1,3)) +
  theme_classic() +
  theme(axis.title.x=element_text(size=10), axis.text.x=element_text(size=10),
        axis.title.y=element_text(size=11), axis.text.y=element_text(size=10),
        legend.title=element_text(size=10), legend.text=element_text(size=10),
        plot.title=element_text(size=11, face='bold.italic', hjust=0.5)) +
  guides(lty='none') +
  
  annotation_custom(ggplotGrob(
    ggplot(subset(simul, Approx_WM), aes(x=time, fill=n_ratio)) +
      geom_ribbon(aes(ymin=0, ymax=qA)) +
      labs(x='Time', y=expression(q^A)) +
      scale_x_continuous(expand=c(0,0), breaks=seq(0,200,100)) +
      scale_y_continuous(expand=c(0,0), limits=c(0,1), breaks=seq(0,1,0.5)) +
      scale_fill_manual(values=greens) +
      scale_linetype_manual(values=c(1,3)) +
      theme_classic() +
      theme(axis.title.x=element_blank(), axis.text.x=element_text(size=6.5),
            axis.title.y=element_text(size=9, angle=0, vjust=0.5), axis.text.y=element_text(size=6.5),
            legend.position='none')),
    xmin=70, xmax=170, ymin=5, ymax=11)
```

## Plot

```{r, fig.width=8, fig.height=4}
ggarrange(Fig2A, Fig2B + theme(axis.title.y=element_blank()),
          ncol=2, common.legend=TRUE, legend='right', labels = LETTERS[1:2], widths=c(0.53,0.47))
```

# The transient disruption of spatial differentiation

```{r, fig.width=8, fig.height=6}
bAw <- bBw <- 0.15
bAm <- bBm <- bBw*1.3

gAw <- gBw <- 0.1
gAm <- gBm <- 0.1

mA <- 1E-2
mB <- 1E-2

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
           "mA"=mA, "mB"=mB)

I0 <- 1E-3

nA <- 100
nratio_vec <- c(0.02,0.1,1,10,50)

simul <- nratio_vec %>% lapply(function(nratio){
  nB <- nratio*nA
  init <- c("SA"=nA-2*I0, "IAw"=I0, "IAm"=I0, "RA"=0,
            "SB"=nB-2*I0, "IBw"=I0, "IBm"=I0, "RB"=0)
  return(simulate(init, times=seq(0,2500,1), parms=parms) %>% cbind('n_ratio'=nratio))
}) %>% bind_rows

S_ratio <- simul[,match(c("time", "Approx_WM", "n_ratio"), colnames(simul))]
S_ratio$S_ratio <- simul$SA/simul$SB*simul$n_ratio

ggplot(simul, aes(x=time, y=lnQ, col=as.factor(n_ratio), lty=Approx_WM)) +
  geom_line(lwd=0.75) +
  labs(x='Time',
       y='Spatial differentiation (log scale)\n',
        col=expression(N^B/N^A)) +
  scale_x_continuous(expand=c(0,0), breaks=c(0,10,50,200,500,1000,2000,5000), trans='sqrt') +
  scale_y_continuous(expand=c(0,0), limits=c(-0.5,0.5), breaks=seq(-0.5,0.5,0.1)) +
  scale_color_manual(values=colorRampPalette(brewer.pal(9,"Greens"))(length(nratio_vec)+3)[-(1:3)]) +
  scale_linetype_manual(values=c(1,3)) +
  theme_minimal() +
  theme(axis.title.x=element_text(size=12), axis.text.x=element_text(size=12),
        axis.title.y=element_text(size=12), axis.text.y=element_text(size=12),
        legend.title=element_text(size=12), legend.text=element_text(size=12)) +
  guides(lty='none')
```

# The transient dynamics of spatial differentiation and the migration–selection balance

```{r}
bAw <- bBw <- bAm <- bBm <- 0.15
gAw <- gBw <- 0.1
gAm <- 0.07
gBm <- 0.04

nA <- 100
nB <- 100

I0w <- 1E-4
q0 <- 1/100

init <- c("SA"=nA-I0w-I0m, "IAw"=I0w, "IAm"=q0/(1-q0)*I0w, "RA"=0,
          "SB"=nB-I0w-I0m, "IBw"=I0w, "IBm"=0, "RB"=0)

tf <- 130
tM <- 50
times <- seq(0,tf,0.1)

theme_main <- theme_classic() +
  theme(axis.title.x=element_blank(), axis.text.x=element_text(size=11),
        axis.title.y=element_text(size=14), axis.text.y=element_text(size=12))

theme_inset <- theme_classic() +
  theme(axis.title.x=element_blank(), axis.text.x=element_text(size=9),
        axis.title.y=element_text(size=10), axis.text.y=element_text(size=9),
        legend.position='none')
```

## No migration

```{r, fig.width=10, fig.height=4}
mA <- mB <- 0

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gAm,
           "mA"=mA, "mB"=mB)

simul <- simulate(init, times[times<=tM], parms, approx.WM=FALSE)

init2 <- simul[nrow(simul),2:9] %>% unlist
init2[["IBm"]] <- q0/(1-q0)*init2[["IBw"]]

simul <- simul[-nrow(simul),] %>% rbind(simulate(init2, times[times>=tM], parms, approx.WM=FALSE))

simul$logit_qB[simul$time<tM] <- simul$lnQ[simul$time<tM] <- NA
simul$time <- simul$time-tM

A <- simul %>% subset(time>=0) %>%
  ggplot(aes(x=time, y=lnQ)) +
  geom_hline(yintercept=0, col='grey60', lwd=0.4, lty="dashed") +
  geom_line(cex=0.7) +
  labs(x="Time", y="Spatial log-differentiation") +
  scale_x_continuous(expand=c(0,0), breaks=seq(0,100,20)) +
  scale_y_continuous(expand=c(0,0), limits=c(-1,6), breaks=seq(-1,6,1)) +
  theme_main +
  
  annotation_custom(ggplotGrob(
    simul[,c("time", "logit_q") %>% sapply(function(x){grep(x, colnames(simul))}) %>% unlist] %>%
      gather(var, freq, -time) %>%
      ggplot(aes(x=time, y=freq, col=var)) +
      geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
      geom_line(cex=0.5) +
      labs(x="Time", y='Variant logit-frequency', col="Pop.") +
      scale_color_manual(values=c("black", '#da6967', '#4c97cc'), labels=c("AB", "A", "B")) +
      scale_x_continuous(expand=c(0,0), breaks=seq(-25,100,25)) +
      scale_y_continuous(expand=c(0,0), limits=c(logit(q0),2), breaks=seq(-6,2,2)) +
      theme_inset +
      theme(legend.title=element_text(size=10), legend.text=element_text(size=10),
            legend.key.height=unit(0.5,'cm'), legend.position="left",
            legend.background=element_rect(colour='black',linewidth=0.2),
            plot.title=element_text(size=15, hjust=0.5))),
    xmin=10, xmax=tf-tM, ymin=2, ymax=6)

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
           "mA"=mA, "mB"=mB)

simul <- simulate(init, times[times<=tM], parms, approx.WM=FALSE)

init2 <- simul[nrow(simul),2:9] %>% unlist
init2[["IBm"]] <- q0/(1-q0)*init2[["IBw"]]

simul <- simul[-nrow(simul),] %>% rbind(simulate(init2, times[times>=tM], parms, approx.WM=FALSE))
simul$logit_qB[simul$time<tM] <- simul$lnQ[simul$time<tM] <- NA
simul$time <- simul$time-tM

B <- simul %>% subset(time>=0) %>%
  ggplot(aes(x=time, y=lnQ)) +
  geom_hline(yintercept=0, col='grey60', lwd=0.4, lty="dashed") +
  geom_line(cex=0.7) +
  labs(x="Time", y="Spatial log-differentiation") +
  scale_x_continuous(expand=c(0,0), breaks=seq(0,100,20)) +
  scale_y_continuous(expand=c(0,0), limits=c(-1,6), breaks=seq(-1,6,1)) +
  theme_main + theme(axis.title.y=element_blank()) +
  
  annotation_custom(ggplotGrob(
    simul[,c("time", "logit_q") %>% sapply(function(x){grep(x, colnames(simul))}) %>% unlist] %>%
      gather(var, freq, -time) %>%
      ggplot(aes(x=time, y=freq, col=var)) +
      geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
      geom_line(cex=0.5) +
      labs(x="Time", y='Variant logit-frequency', col="Population") +
      scale_color_manual(values=c("black", '#da6967', '#4c97cc'), labels=c("AB", "A", "B")) +
      scale_x_continuous(expand=c(0,0), breaks=seq(-25,100,25)) +
      scale_y_continuous(expand=c(0,0), limits=c(logit(q0),2), breaks=seq(-6,2,2)) +
      theme_inset + theme(legend.position='none')),
    xmin=20, xmax=tf-tM, ymin=1.5, ymax=6)

(AB <- plot_grid(A, B, ncol=2, labels=LETTERS[1:2], align='hv'))
```

## Low levels of migration

```{r, fig.width=10, fig.height=4}
mA <- mB <- 1E-2

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gAm,
           "mA"=mA, "mB"=mB)

simul <- simulate2(init, times, parms, tM, qB0=q0) %>% subset(Approx_WM==FALSE)
simul$qB[simul$time<tM] <- simul$lnQ[simul$time<tM] <- simul$lnQ_QE[simul$time<tM] <- NA
simul$time <- simul$time-tM

C <- simul %>% subset(time>=0) %>%
  ggplot(aes(x=time, y=lnQ)) +
  geom_hline(yintercept=0, col='grey60', lwd=0.4, lty="dashed") +
  geom_line(cex=0.7) +
  labs(x="Time", y="Spatial log-differentiation") +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0), limits=c(-1,6), breaks=seq(-1,6,1)) +
  theme_main +
  
  annotation_custom(ggplotGrob(
    simul[,c("time", "logit_q") %>% sapply(function(x){grep(x, colnames(simul))}) %>% unlist] %>%
      gather(var, freq, -time) %>%
      ggplot(aes(x=time, y=freq, col=var)) +
      geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
      geom_line(cex=0.5) +
      labs(x="Time", y='Variant logit-frequency', col="Population") +
      scale_color_manual(values=c("black", '#da6967', '#4c97cc'), labels=c("AB", "A", "B")) +
      scale_x_continuous(expand=c(0,0)) +
      scale_y_continuous(expand=c(0,0), limits=c(logit(q0),2), breaks=seq(-6,2,2)) +
      theme_inset),
    xmin=20, xmax=tf-tM, ymin=1.5, ymax=6)

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
           "mA"=mA, "mB"=mB)

simul <- simulate2(init, times, parms, tM, qB0=q0) %>% subset(Approx_WM==FALSE)
simul$qB[simul$time<tM] <- simul$lnQ[simul$time<tM] <- simul$lnQ_QE[simul$time<tM] <- NA
simul$time <- simul$time-tM

D <- simul %>% subset(time>=0) %>%
  ggplot(aes(x=time, y=lnQ)) +
  geom_hline(yintercept=0, col='grey60', lwd=0.4, lty="dashed") +
  geom_line(cex=0.7) +
  labs(x="Time", y="Spatial log-differentiation") +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0), limits=c(-1,6), breaks=seq(-1,6,1)) +
  theme_main + theme(axis.title.y=element_blank()) +
  
  annotation_custom(ggplotGrob(
    simul[,c("time", "logit_q") %>% sapply(function(x){grep(x, colnames(simul))}) %>% unlist] %>%
      gather(var, freq, -time) %>%
      ggplot(aes(x=time, y=freq, col=var)) +
      geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
      geom_line(cex=0.5) +
      labs(x="Time", y='Variant logit-frequency', col="Population") +
      scale_color_manual(values=c("black", '#da6967', '#4c97cc'), labels=c("AB", "A", "B")) +
      scale_x_continuous(expand=c(0,0)) +
      scale_y_continuous(expand=c(0,0), limits=c(logit(q0),2), breaks=seq(-6,2,2)) +
      theme_inset),
    xmin=20, xmax=tf-tM, ymin=1.5, ymax=6)

(CD <- plot_grid(C, D, ncol=2, labels=LETTERS[3:4], align='hv'))
```

## Higher levels of migration

```{r, fig.width=10, fig.height=4}
mA <- mB <- 0.1

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gAm,
           "mA"=mA, "mB"=mB)

simul <- simulate2(init, times, parms, tM, qB0=q0) %>% subset(Approx_WM==FALSE)
simul$qB[simul$time<tM] <- simul$lnQ[simul$time<tM] <- simul$lnQ_QE[simul$time<tM] <- NA
simul$time <- simul$time-tM

E <- subset(simul, time>=0) %>%
  ggplot(aes(x=time, y=lnQ)) +
  geom_hline(yintercept=0, col='grey60', lwd=0.4, lty="dashed") +
  geom_line(cex=0.7) +
  labs(x="Time", y="Spatial log-differentiation") +
  scale_linetype_manual(values = 1:2) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0), limits=c(-1,6), breaks=seq(-1,6,1)) +
  theme_main + theme(axis.title.x=element_text(size=14)) + 
  
  annotation_custom(ggplotGrob(
    simul[,c("time", "logit_q") %>% sapply(function(x){grep(x, colnames(simul))}) %>% unlist] %>%
      gather(var, freq, -time) %>%
      ggplot(aes(x=time, y=freq, col=var)) +
      geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
      geom_line(cex=0.5) +
      labs(x="Time", y='Variant logit-frequency', col="Population") +
      scale_color_manual(values=c("black", '#da6967', '#4c97cc'), labels=c("AB", "A", "B")) +
      scale_x_continuous(expand=c(0,0)) +
      scale_y_continuous(expand=c(0,0), limits=c(logit(q0),2), breaks=seq(-6,2,2)) +
      theme_inset),
    xmin=20, xmax=tf-tM, ymin=1.5, ymax=6)

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
           "mA"=mA, "mB"=mB)

simul <- simulate2(init, times, parms, tM, qB0=q0) %>% subset(Approx_WM==FALSE)
simul$qB[simul$time<tM] <- simul$lnQ[simul$time<tM] <- simul$lnQ_QE[simul$time<tM] <- NA
simul$time <- simul$time-tM

G <- subset(simul, time>=0) %>%
  ggplot(aes(x=time, y=lnQ)) +
  geom_hline(yintercept=0, col='grey60', lwd=0.4, lty="dashed") +
  geom_line(cex=0.7) +
  labs(x="Time", y="Spatial log-differentiation") +
  scale_linetype_manual(values = 1:2) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0), limits=c(-1,6), breaks=seq(-1,6,1)) +
  theme_main + theme(axis.title.x=element_text(size=14), axis.title.y=element_blank()) +

  annotation_custom(ggplotGrob(
    simul[,c("time", "logit_q") %>% sapply(function(x){grep(x, colnames(simul))}) %>% unlist] %>%
      gather(var, freq, -time) %>%
      ggplot(aes(x=time, y=freq, col=var)) +
      geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
      geom_line(cex=0.5) +
      labs(x="Time", y='Variant logit-frequency', col="Population") +
      scale_color_manual(values=c("black", '#da6967', '#4c97cc'), labels=c("AB", "A", "B")) +
      scale_x_continuous(expand=c(0,0)) +
      scale_y_continuous(expand=c(0,0), limits=c(logit(q0),2), breaks=seq(-6,2,2)) +
      theme_inset),
    xmin=20, xmax=tf-tM, ymin=1.5, ymax=6)

(EF <- plot_grid(E, G, ncol=2, labels=LETTERS[5:6], align='hv'))
```

## Plot assembly

```{r, fig.width=9, fig.height=10}
plot_grid(plot_grid(ggdraw()+draw_label(expression('\U1D4AE '^A~'='~'\U1D4AE '^B), size=22, hjust=0),
                    ggdraw()+draw_label(expression('\U1D4AE '^A~'<'~'\U1D4AE '^B), size=22, hjust=0),
                    ggdraw(), ncol=3, rel_widths=c(0.475,0.475,0.05)),
          plot_grid(AB, ggdraw()+draw_label(expression(bold('No migration')), angle=-90, size=16),
                    ncol=2, rel_widths=c(0.95,0.05)),
          plot_grid(CD, ggdraw()+draw_label(expression(bold('Migration prob.'~'='~10^{-2})), size=16, angle=-90),
                    ncol=2, rel_widths=c(0.95,0.05)),
          plot_grid(EF, ggdraw()+draw_label(expression(bold('Migration prob.'~'='~0.1)), size=16, angle=-90),
                    ncol=2, rel_widths=c(0.95,0.05)),
          ncol=1, rel_heights=c(0.08,0.46,0.46,0.46))

# ggsave(width=9, height=10, dpi=600, 'Differentiation_Delta_gamma_V2.png')
```

# Long-term spatial differentiation and local adaptation

```{r}
bAw <- bBw <- bAm <- bBm <- 0.15
gAw <- gBw <- 0.1
gBm <- 0.05
w <- 0.01

gAm_vec <- seq(0.05,0.15,length.out=300)
m_vec <- seq(0,0.075,length.out=300)

times <- seq(0,1E6,5E5)

X <- expand.grid(gAm_vec, m_vec) %>% setNames(c("gAm", "m"))
X$diff_Delta_gamma <- -(X$gAm-gAw)+(gBm-gBw)
X$lnQ <- NA
X$lnQ2 <- NA
```

```{r, eval=FALSE}
nA <- 100
nB <- 100

for(i in 1:nrow(X)){
  
  gAm <- X$gAm[i]
  m <- X$m[i]
  
  if(gAm<gAw){
    init <- c("SA"=nA*gAm/bAm,"IAw"=0,
              "IAm"=nA*(bAm-gAm)*w/(bAm*(gAm+w)),"RA"=nA*(1-gAm/bAm-(bAm-gAm)*w/(bAm*(gAm+w))),
              "SB"=nB*gBm/bBm,"IBw"=0,
              "IBm"=nB*(bBm-gBm)*w/(bBm*(gBm+w)),"RB"=nB*(1-gBm/bBm-(bBm-gBm)*w/(bBm*(gBm+w))))
  }else{
    init <- c("SA"=nA*gAw/bAw,"IAw"=nA*(bAw-gAw)*w/(bAw*(gAw+w)),
              "IAm"=0,"RA"=nA*(1-gAw/bAw-(bAw-gAw)*w/(bAw*(gAw+w))),
              "SB"=nB*gBm/bBm,"IBw"=0,
              "IBm"=nB*(bBm-gBm)*w/(bBm*(gBm+w)),"RB"=nB*(1-gBm/bBm-(bBm-gBm)*w/(bBm*(gBm+w))))
  }
  
  parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bAm,
             "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
             "mA"=m, "mB"=m, "w"=w)
  
  yend <- lsoda(y=init, times=times, func=SIR_metapop, parms=parms)[length(times),]
  
  qA_end <- yend[["IAm"]]/(yend[["IAw"]]+yend[["IAm"]])
  qB_end <- yend[["IBm"]]/(yend[["IBw"]]+yend[["IBm"]])
  X$lnQ[i] <- logit(qA_end)-logit(qB_end)
  
  f <- (yend[["IAw"]]+yend[["IAm"]])/(yend[["IAw"]]+yend[["IAm"]]+yend[["IBw"]]+yend[["IBm"]])
  n_ratio <- nB/nA
  
  X$Q_approx[i] <- 1+X$diff_Delta_gamma[i]/
    ((m*bAw/nA+m*bBw/nB)*(qB_end/qA_end*(1-f)/f*yend[["SA"]]+(1-qA_end)/(1-qB_end)*f/(1-f)*yend[["SB"]]))
}

# saveRDS(X, file="long_term_diff.RDS")
```

```{r, eval=FALSE}
nA <- 100
nB <- 1000

for(i in 1:nrow(X)){
  
  gAm <- X$gAm[i]
  m <- X$m[i]
  
  if(gAm<gAw){
    init <- c("SA"=nA*gAm/bAm,"IAw"=0,
              "IAm"=nA*(bAm-gAm)*w/(bAm*(gAm+w)),"RA"=nA*(1-gAm/bAm-(bAm-gAm)*w/(bAm*(gAm+w))),
              "SB"=nB*gBm/bBm,"IBw"=0,
              "IBm"=nB*(bBm-gBm)*w/(bBm*(gBm+w)),"RB"=nB*(1-gBm/bBm-(bBm-gBm)*w/(bBm*(gBm+w))))
  }else{
    init <- c("SA"=nA*gAw/bAw,"IAw"=nA*(bAw-gAw)*w/(bAw*(gAw+w)),
              "IAm"=0,"RA"=nA*(1-gAw/bAw-(bAw-gAw)*w/(bAw*(gAw+w))),
              "SB"=nB*gBm/bBm,"IBw"=0,
              "IBm"=nB*(bBm-gBm)*w/(bBm*(gBm+w)),"RB"=nB*(1-gBm/bBm-(bBm-gBm)*w/(bBm*(gBm+w))))
  }
  
  parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bAm,
             "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
             "mA"=m, "mB"=m, "w"=w)
  
  yend <- lsoda(y=init, times=times, func=SIR_metapop, parms=parms)[length(times),]
  
  qA_end <- yend[["IAm"]]/(yend[["IAw"]]+yend[["IAm"]])
  qB_end <- yend[["IBm"]]/(yend[["IBw"]]+yend[["IBm"]])
  X$lnQ[i] <- logit(qA_end)-logit(qB_end)
  
  f <- (yend[["IAw"]]+yend[["IAm"]])/(yend[["IAw"]]+yend[["IAm"]]+yend[["IBw"]]+yend[["IBm"]])
  n_ratio <- nB/nA
  
  X$Q_approx[i] <- 1+X$diff_Delta_gamma[i]/
    ((m*bAw/nA+m*bBw/nB)*(qB_end/qA_end*(1-f)/f*yend[["SA"]]+(1-qA_end)/(1-qB_end)*f/(1-f)*yend[["SB"]]))
}

# saveRDS(X, file="long_term_diff2.RDS")
```

```{r, fig.width=7.5, fig.height=3.5}
plot_longterm_diff <- function(data, N_ratio=1){
  
  # Hatched background
  rect_df <- data.frame(xmin=-(gAm_vec[which(gAm_vec>gAw)[1]]-gAw)+gBm-gBw,
                        xmax=0, ymin=0, ymax=max(m_vec), fill='fill')
  return(
    data %>%
      ggplot(aes(x=diff_Delta_gamma, y=m, fill=lnQ)) +
      geom_tile() +
      scale_fill_viridis_c(limits=c(-15,-4), name="ln(\U1D4AC)", guide=guide_colorbar(order=2)) +
      ggnewscale::new_scale_fill() +
      geom_rect_pattern(data=rect_df, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=fill),
                        inherit.aes=FALSE, pattern="stripe", pattern_angle=45, pattern_density=0.5,
                        pattern_spacing=0.15, pattern_size=0.15, size=0.15, pattern_fill="grey50", color="black") +
      scale_fill_manual(name="", values=c('fill'="grey50"), labels=c('fill'="Variant adapted\nin both pop."),
                        guide=guide_legend(order=1,override.aes=list(pattern_spacing=0.05,
                                                                     pattern_size=0.2,
                                                                     size=5))) +
      labs(title=bquote(N^B/N^A==.(N_ratio)),
           x=expression('\U1D4AE '^A-'\U1D4AE '^B), y=expression(paste('Migration ',mu))) +
      scale_x_continuous(expand=c(0,0), limits=c(-0.1,0)) +
      scale_y_continuous(expand=c(0,0)) +
      theme(axis.title.x=element_text(size=12), axis.title.y=element_text(size=12),
            legend.title=element_text(size=14), plot.title=element_text(hjust=0.5))
  )
}

ggarrange(
  plot_longterm_diff(readRDS("long_term_diff.RDS")),
  plot_longterm_diff(readRDS("long_term_diff2.RDS"), N_ratio=10),
  ncol=2, common.legend=TRUE, legend='right', labels=LETTERS[1:2])

# ggsave(width=7.5, height=3.5, dpi=500, 'Long_term_lnQ_V2.png')
```

# Reduced mobility of infected hosts

```{r}
nA <- 100
nB <- 100

q0 <- 1/100

I0w <- 1E-7
I0m <- q0/(1-q0)*I0w

init <- c("SA"=nA-I0w-I0m, "IAw"=I0w, "IAm"=I0m, "RA"=0,
          "SB"=nB-I0w, "IBw"=I0w, "IBm"=0, "RB"=0)

tf <- 200
tM <- 50
times <- seq(0,tf,1)

bAw <- bBw <- 0.15
bAm <- bBm <- bBw*1.3

gAw <- gBw <- 0.1
gAm <- gBm <- 0.1

mA <- mB <- 0.1

parms <- c("bAw"=bAw, "bBw"=bBw, "bAm"=bAm, "bBm"=bBm,
           "gAw"=gAw, "gBw"=gBw, "gAm"=gAm, "gBm"=gBm,
           "mA"=mA, "mB"=mB)

z_vec <- seq(0,1,0.25)

simul <- z_vec %>% lapply(FUN=function(z){
  parms[["z"]] <- z
  return(data.frame("z"=z, simulate2(init=init, times=times, parms=parms, tM=tM, qB0=q0)))
  }) %>% bind_rows %>% subset(Approx_WM==FALSE)

simul$time <- simul$time-tM

colors <- magma(length(z_vec)+1)[-(length(z_vec)+1)]

(Fig_zeta1 <- ggarrange(
  data.frame("z"=simul$z, "time"=simul$time, "logit_q"=c(simul$logit_qA,simul$logit_qB),
             "pop"=rep(c("Population A","Population B"), each=nrow(simul))) %>%
    ggplot(aes(x=time, y=logit_q, col=as.factor(z))) +
    facet_wrap(~pop, ncol=1) +
    geom_rect(aes(xmin=-Inf, xmax=0, ymin=-Inf, ymax=+Inf), fill='grey93', col=NA) +
    geom_line(aes(order=(z==0)), size=0.35) +
    scale_color_manual(values=colors) +
    labs(x='Time', y='Logit-frequency of the variant\n', col=expression(paste(zeta))) +
    scale_x_continuous(limits=c(-tM,75), breaks=seq(-50,75,25), expand=c(0,0)) +
    scale_y_continuous(expand=c(0,0), limits=c(logit(q0),1), breaks=seq(-4,1,1)) +
    theme_bw() +
    theme(axis.title.x=element_text(size=9), axis.text.x=element_text(size=9),
          axis.title.y=element_text(size=10), axis.text.y=element_text(size=8)),

  ggplot(simul, aes(x=time, y=lnQ, col=as.factor(z))) +
    geom_hline(yintercept=0, lwd=0.3, lty='dotted') +
    geom_line(aes(order=(z==0))) +
    labs(x='Time', y='Spatial differentiation (log scale)\n', col=expression(paste(zeta))) +
    scale_x_continuous(limits=c(0,tf-tM+5), breaks=seq(-50,tf-tM,50), expand=c(0,0)) +
    scale_color_manual(values=colors) +
    annotate(geom="text", x=140, y=2.1, label="N^B/N^A==1", parse=TRUE, size=3, hjust=1) +
    annotate(geom="text", x=140, y=1.9, label="mu^B/mu^A==1", parse=TRUE, size=3, hjust=1) +
    theme_bw() +
    theme(axis.title.x=element_text(size=9), axis.text.x=element_text(size=9),
          axis.title.y=element_text(size=10), axis.text.y=element_text(size=9)),
  
  ncol=2, labels=paste0("A", 1:2), common.legend=TRUE, legend="top")) # 3.5 x 6
```

```{r}
simul <- data.frame()

n_ratio_vec <- c(0.1,1,10)
m_ratio_vec <- c(0.1,1,2)

for(n_ratio in n_ratio_vec){
  
  nB <- n_ratio*nA
  init <- c("SA"=nA-I0w-I0m, "IAw"=I0w, "IAm"=I0m, "RA"=0,
            "SB"=nB-I0w, "IBw"=I0w, "IBm"=0, "RB"=0)
  
  for(m_ratio in m_ratio_vec){
    
    parms[["mB"]] <- mA*m_ratio
    
    simul <- simul %>% rbind(
      z_vec %>% lapply(FUN=function(z){
        parms[["z"]] <- z
        return(data.frame("z"=z, simulate2(init=init, times=times, parms=parms, tM=tM, qB0=q0)))
        }) %>% bind_rows %>% subset(Approx_WM==FALSE) %>% cbind("n_ratio"=n_ratio, "m_ratio"=m_ratio)
      ) 
  }
}

simul$time <- simul$time-tM

(Fig_zeta2 <- simul %>%
  mutate(
    n_ratio=factor(n_ratio, levels=n_ratio_vec,
                   labels=sapply(n_ratio_vec, function(x){bquote(N^B/N^A==.(x))})),
    m_ratio=factor(m_ratio, levels=m_ratio_vec,
                   labels=sapply(n_ratio_vec, function(x){bquote(mu^B/mu^A==.(x))}))
    ) %>%
  ggplot(aes(x=time, y=lnQ, col=as.factor(z))) +
  facet_grid(m_ratio~n_ratio, labeller=label_parsed) +
    geom_hline(yintercept=0, lwd=0.3, lty='dotted') +
    geom_line(aes(order=(z==0))) +
    labs(x='Time', y='Spatial differentiation (log scale)\n', col=expression(paste(zeta))) +
    scale_x_continuous(limits=c(0,tf-tM+5), breaks=seq(-50,tf-tM,50), expand=c(0,0)) +
    scale_color_manual(values=colors) +
    theme_bw() +
    theme(axis.title.x=element_text(size=9), axis.text.x=element_text(size=9),
          axis.title.y=element_text(size=10), axis.text.y=element_text(size=9)))
```

```{r}
plot_grid(Fig_zeta1, Fig_zeta2, labels=c("", "B"), ncol=1)
```
